A Java Framework For Embarrassingly Parallel Problems

### 2013-12-10 paper/paper.tex
More paper changes. Tables etc

Ka/Ks Calculator is not the only embarrassingly parallel program that has the
potential to take several days to run on a data set. In fact, there are several
embarrassingly parallel programs in the biological sciences. Basic Local
Alignment Search Tool, or BLAST\fotenote{blast.ncbi.nlm.nih.gov}, is an
algorithm that compares many aspects of DNA sequences, like the nucleotides or
amino acid sequences. In addition, the user can also compare to the sequences to
a database of sequences to identity known subservience's. There are many
implementations of BLAST, some of them already have parallel implementations.
Many applications however do not have parallel processing capabilities. A class
of problems that organize large amounts of data into similar groups called
clusters. Two applications, octupus\footnote{octupus.sourceforge.com} and
USEARCH\footnote{https://www.drive5.com/usearch}, are examples of programs in
this class that do not have parallel versions. With several applications that
would benefit from parallelism in this field, it was obvious it would be
desirable to develop a framework for the parallelization task.

@@

A version of octupus was also run through the framework and was tested against
two different sized data sets. Both sets were first clustered at 90\% similarity
and the resulting clusters were used as the input for the framework's run, which
is for 97\% similarity. The framework in this case was just run at four and
eight threads and Figure \ref{fig:octugraph} shows that the running the
framework once again provides a benefit for the user.

\begin{figure}
{\resizebox{5.2in}{!}{\includegraphics{experiment/outoctu.png}}}
\caption{Runtimes for the Ka/Ks Calculator though the framework}
\label{fig:octugraph}
\end{figure}

>>>>

Ka/Ks Calculator is not the only embarrassingly parallel program that has the
potential to take several days to run on a data set. In fact, there are several
embarrassingly parallel programs in the biological sciences. Basic Local
Alignment Search Tool, or BLAST\footnote{blast.ncbi.nlm.nih.gov}, is an
algorithm that compares many aspects of DNA sequences, like the nucleotides or
amino acid sequences. In addition, the user can also compare to the sequences to
a database of sequences to identity known subservience's. There are many
implementations of BLAST, some of them already have parallel implementations.
Many applications however do not have parallel processing capabilities. A class
of problems that organize large amounts of data into similar groups called
clusters. Two applications, octupus\footnote{octupus.sourceforge.com} and
USEARCH\footnote{https://www.drive5.com/usearch}, are examples of programs in
this class that do not have parallel versions. With several applications that
would benefit from parallelism in this field, it was obvious it would be
desirable to develop a framework for the parallelization task.

@@

A version of octupus was also run through the framework and was tested against
two different sized data sets. Both sets were first clustered at 90\%, 95\% and
97\% minimum similarity. The runtime for those tests can be seen in table
\ref{tab:oct1}. The resulting clusters from the 90\% similarity run were used as
the input for the framework's runs, which were all set for 97\% similarity. 
The framework in this case was just run at six, eight and ten workers. Table 
\ref{tab:octu2} shows that it is slightly faster to cluster first at 90\% and 
then run the resulting clusters through the framework at a higher similarity.

\begin{table}
    \begin{tabular}{|l|l|l|}
        \hline
        Number of Sequences & Minimum Similarity & Time(s) \\ \hline
        1500                & 90\%               & 99      \\
        ~                   & 95\%               & 115     \\
        ~                   & 97\%               & 131     \\ \hline
        6500                & 90\%               & 425     \\
        ~                   & 95\%               & 496     \\
        ~                   & 97\%               & 603     \\ \hline
    \end{tabular}
    \label{tab:octu1}
\end{table}

\begin{table}
    \begin{tabular}{|l|l|l|}
        \hline
        Number of Sequences & Number of Workers & Time(s) \\ \hline
        1500                & 6                 & 34      \\
        ~                   & 8                 & 27      \\
        ~                   & 10                & 27      \\ \hline
        6500                & 6                 & 127     \\
        ~                   & 8                 & 127     \\
        ~                   & 10                & 123     \\ \hline
    \end{tabular}
    \label{tab:octu2}
\end{table}
